Crossing Origins by Crossing Formats

### 2012-11-27 main.tex
Updated section on PDF

%3 pages
\section{Crossing origins by crossing formats}

% Describe how the attacks lead to Same-Origin Policy bypass.

@@

In this section we give concrete examples of the attacks described in the
previous section, using the PDF format as the running example.

\subsection{Concept}

@@

The PDF format is a widely used, text based format with powerful capabilities.
It is capable of displaying text, rendering graphics, animation and dynamic
content. It is a container format which allows for embedding of other files
and formats. This and the fact that it is designed to be robust against errors,
provides us with a format that is powerful, flexible and ideal for creating
malicious polyglots.

@@

% - Error tolerant parsing
% - Execute JavaScript
% - Embed flash files
% - Launch commands

@@

For communication with the containing page, one can either use
hostContainer.postMessage(info) or app.launchURL('\#'+info). The latter has the
advantage that it works in Firefox which seems to not handle the postMessage
very well.

@@

var s = 'user string<\\/script>';

>>>>

%3 pages
\section{Crossing origins by crossing formats}
\label{sec:cross}
% Describe how the attacks lead to Same-Origin Policy bypass.

@@

In this section we give concrete examples of the attacks described in 
Section~\ref{sec:cross}, using the PDF format as the running example. We begin
by detailing how the design decision made in the PDF-standard make it ideal 
for creating malicious polyglots.

%\subsection{Concept}

@@

The \emph{Portable Document Format} (PDF) is a widely used document format capable of displaying text, 
rendering graphics, scripting, animation and other dynamic content.

\subsubsection{File structure}
% header
% objects
% cross-reference
% trailer

According to the PDF specification \todo{cite specification} a PDF document consists of
a header, several objects, a cross-reference section and a trailer.

\paragraph{Header}
The header consists of the string "\%PDF-" followed by a version number. 
According to the PDF reference implementation notes, the header can occur 
anywhere within the first 1024 bytes of the file. The version is denoted by a 
major and a minor version number of the form "M.m". Because the version can 
be specified elsewhere, the version number is not required to be part of the 
header.

\paragraph{Objects}
Objects can be direct or indirect, the difference being that indirect objects 
has labels which are used for referring to the object from another object. Object labels 
There are eight basic types of basic objects; booleans, integers, strings, names, 
arrays, dictionaries, streams and the null object. For the intents and purposes 
of this article, we will focus on the string, dictionary and stream objects.

% dictionary objects
% string objects
% stream objects

String objects are delimited by the "(" and ")" characters.

Dictionary objects are delimited by the "<<" and ">>" tokens.

Stream objects

\paragraph{Cross-reference}
The cross-reference section is a record of the location of indirect objects 
within a file. The location is specified as the byte offset from the beginning 
of the file. The cross-reference section is opened with the "xref" keyword,
followed by one record for each indirect object. The cross-reference section 
will be reconstructed if missing and can therefore be omitted.

\paragraph{Trailer}
The trailer is composed of a trailer dictionary, a pointer to the cross-reference 
section and an end-of-file marker. The trailer dictionary is introduced by the 
"trailer" keyword. It contains among other things a reference to the root of the 
object tree. The "startxref" keyword is followed by the number of bytes from 
the beginning of the file to the first cross-reference section. As with the 
cross-reference section, it can be omitted. The string "\%\%EOF" marks the 
end-of-file, but can be omitted.

% actions?

\begin{lstlisting}[label=lst:sample-pdf, caption=HTML / E4X Polyglot]
%PDF-1.7
1 0 obj<</Type/Catalog/Pages<<>>>>
endobj
2 0 obj<</Length 13>>
stream
Sample Author
endstream
endobj
xref
0000000000 65536 f
0000000009 00000 n
0000000057 00000 n
trailer<</Root 1 0 R/Info<</Author 2 0 obj>>>>
startxref
127
%%EOF
\end{lstlisting}

\subsubsection{PDF-based polyglots}
Several design choices in the PDF specification make the format particularly 
suitable for making polyglots. It is a container format in the sense that it allows 
embedding of files and resources. Therefore a PDF file can, by design, contain foreign syntax that 
could interfere with its syntax and hence the parsing is designed to be error 
tolerant. Another motivation for this design choice is that PDF files can be 
incrementally updated, by appending updated versions of existing objects to 
the end of the file. In fact, 
%and formats. This and the fact that it is designed to be robust against errors,
%provides us with a format that is powerful, flexible and ideal for creating
%malicious polyglots.

@@

% - Error tolerant parsing
% - Backward compatibility
% - Container format
% - Execute JavaScript
% - Embed flash files
% - Launch commands

@@

For communication with the containing page, one can either use
hostContainer.postMessage(info) or app.launchURL('\#' + info). The latter has the
advantage that it works in Firefox which seems to not handle the postMessage
very well.

@@

var s = 'user string</script>';
